Transparent Standby for Low-Power, Resource-Constrained Embedded Systems

### 2018-02-17 LCTES/body.tex
(>) formatting

{\linespread{1}
\begin{figure}[t]
\begin{minipage}[t]{0.49\linewidth}
\begin{lstlisting}[xrightmargin=0.5cm]
while (1) {
  delay(1000);
  int v =
    analogRead(A0);
  radioWrite(v);
}
\end{lstlisting}
\centering\small{[a] Version in Arduino}
\end{minipage}
%
\begin{minipage}[t]{0.49\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=-0.2cm]
loop do
  await 1s;
  var int v =
    await AnalogRead(A0);
  await RadioWrite(v);
end
\end{lstlisting}
\centering\small{[b] Version in \CEU}
\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ Sequence of I/O operations running in a loop.
\label{lst.direct}
}
\end{figure}
}

@@

\begin{figure}[t]
\begin{minipage}[t]{0.49\linewidth}
\begin{lstlisting}[xrightmargin=0.5cm]
uint32_t prv =
  millis();
while (1) {
  if (radioAvail()) {
    break;
  }
  uint32_t cur =
    millis();
  if (cur>prv+1000) {
    prv = cur;
    int v =
      analogRead(A0);
    radioWrite(v);
  }
}
\end{lstlisting}
\centering\small{[a] Version in Arduino}
\end{minipage}
%
\begin{minipage}[t]{0.49\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=-0.2cm]
par/or do
  await RadioAvail();
with
  loop do
    await 1s;
    var int v =
      await AnalogRead(A0);
    await RadioWrite(v);
  end
end

@@

\section{Transparent Standby Mechanisms}
\label{sec.ceu}

\begin{figure}[t]
\begin{lstlisting}[numbers=left]
output (int pin) ANALOG_REQUEST do
    <...>   // port manipulation to start the conversion
end

async/isr ADC_vect_num do
    var int value = <...>; // port with the value read
    emit ANALOG_DONE(value);
end

code AnalogRead (var int pin) -> int do
    {PM_SET(PM_ANALOG, 1);}
    do finalize with
        {PM_SET(PM_ANALOG, 0);}
    end

@@

\begin{figure}[t]
\begin{lstlisting}[numbers=left]
#define PM_GET(peripheral)       bitRead(pm, peripheral)
#define PM_SET(peripheral,state) bitWrite(pm, peripheral, state);

>>>>

{\linespread{1}
\begin{figure}[t]
\begin{minipage}[t]{0.49\linewidth}
\begin{lstlisting}[xrightmargin=0.5cm]
while (1) {
  delay(1000);
  int v =
    analogRead();
  radioWrite(v);
}
\end{lstlisting}
\centering\small{[a] Version in Arduino}
\end{minipage}
%
\begin{minipage}[t]{0.49\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=-0.2cm]
loop do
  await 1s;
  var int v =
    await AnalogRead();
  await RadioWrite(v);
end
\end{lstlisting}
\centering\small{[b] Version in \CEU}
\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ Sequence of I/O operations running in a loop.
\label{lst.direct}
}
\end{figure}
}

@@

\begin{figure}[t]
\begin{minipage}[t]{0.49\linewidth}
\begin{lstlisting}[xrightmargin=0.5cm]
uint32_t prv =
  millis();
while (1) {
  if (radioAvail()) {
    break;
  }
  uint32_t cur =
    millis();
  if (cur>prv+1000) {
    prv = cur;
    int v =
      analogRead();
    radioWrite(v);
  }
}
\end{lstlisting}
\centering\small{[a] Version in Arduino}
\end{minipage}
%
\begin{minipage}[t]{0.49\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=-0.2cm]
par/or do
  await RadioAvail();
with
  loop do
    await 1s;
    var int v =
      await AnalogRead();
    await RadioWrite(v);
  end
end

@@

\section{Transparent Standby Mechanisms}
\label{sec.standby}

In order to empower the example in Figure~\ref{lst.inversion}.b with automatic
standby, we had to make some modifications and extensions to \CEU:
%
\begin{itemize}
\item Modify the runtime of \CEU to be interrupt driven and to put the
      microcontroller in standby after each reaction to the environment.
\item Provide operations for the drivers to indicate which interrupts can awake
      the program.
\item Support ISRs in \CEU to generate input events to the program.
\end{itemize}

Figure~\ref{lst.adc} shows the driver for .

\begin{figure}[t]
\begin{lstlisting}[numbers=left]
// Exposed driver functionality

output void ANALOG_REQUEST;    // low-level request
input  int  ANALOG_DONE;       // low-level response
code AnalogRead (void) -> int; // high-level abstraction

// Driver implementation

output void ANALOG_REQUEST do
  <...>   // port manipulation to start the conversion
end

async/isr ADC_vect_num do
  var int value = <...>; // register with the value read
  emit ANALOG_DONE(value);
end

code AnalogRead (void) -> int do
  {PM_SET(PM_ANALOG, 1);}
  do finalize with
      {PM_SET(PM_ANALOG, 0);}
  end

@@

\begin{figure}[t]
\begin{lstlisting}[numbers=left]
#define PM_GET(peripheral) \
        bitRead(pm, peripheral)
#define PM_SET(peripheral,state) \
        bitWrite(pm, peripheral, state)
